---
title: "Benthic Data Pin"
output: html_document
resource_files:
- config.yml
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(sf)
library(pool)
library(geojsonsf)
library(DBI)
library(pins)
library(config)
library(lubridate)

# get configuration settings
conn <- config::get("connectionSettings")

# use API key to register board
board_register_rsconnect(key = conn$CONNECT_API_KEY,  #Sys.getenv("CONNECT_API_KEY"),
                          server = conn$CONNECT_SERVER)#Sys.getenv("CONNECT_SERVER"))

source('helperFunctions/VSCI_metrics_GENUS.R')
source('helperFunctions/VCPMI_metrics_GENUS.R')

```

```{r connect to ODS_test}
## Connect to ODS_test

# establish db connection locally
#pool <- dbPool(
#  drv = odbc::odbc(),
#  Driver = "SQL Server",  # note the space between SQL and Server ( how MS named driver)
#  Server= "WSQ04151,50000",
#  dbname = "ODS_test"
#)

# establish db connection on server
pool <- dbPool(
  drv = odbc::odbc(),
  Driver = "SQLServer",   # note the LACK OF space between SQL and Server ( how RStudio named driver)
  Server = "DEQ-SQLODS-PROD,50000",
  dbname = "ODS",
  UID = conn$UID_prod,# use config file         #Sys.getenv("userid"), # plug this in on server
  PWD = conn$PWD_prod, # use config file         #Sys.getenv("pwd") # plug this in on server
  trusted_connection = "yes"
)

# con <- dbConnect(odbc::odbc(),.connection_string = 'driver={SQL Server Native Client 11.0};server={DEQ-SQLODS-PROD,50000};database={ODS};trusted_connection=yes')

```

Pull Genus Master Taxa list.

```{r}
masterTaxaGenus <- pool %>% tbl("Edas_Benthic_Master_Taxa_View") %>%
  as_tibble() %>%
  # make columns match expected format
  rename('Phylum' = 'PHYLUM_NAME',
         'Class' = 'CLASS_NAME',
         'Subclass' = 'SUBCLASS_NAME',
         'Order' = 'ORDER_NAME',
         'Suborder' = 'SUBORDER_NAME',
         'Superfamily' = 'SUPERFAMILY_NAME',
         'Family' = 'FAMILY_NAME',
         'Subfamily' = 'SUBFAMILY_NAME',
         'Tribe' = 'TRIBE_NAME',
          'Genus' = 'GENUS_NAME',
          'Species' = 'SPECIES_NAME',
          "Final VA Family ID" =  "WBMT_FINAL_FAMILY_ID",
          "FinalID" = "WBMT_FINAL_ID",
          "TolVal" = "WBMT_TOLERANCE_VALUE",
          "FFG" =   "FEEDING_GROUP", 
          "Habit" = "HABIT", 
          "FamFFG" =  "FAMILY_FEEDING_GROUP",
          "FamTolVal" = "WBMT_FAM_TOLERANCE_VALUE",
          "FamHabit" ="FAMILY_HABIT") %>% 
  dplyr::select(Phylum, Class, Subclass, Order, Suborder, Superfamily, Family, Subfamily, Tribe, 
                Genus, Species, `Final VA Family ID`, FinalID, TolVal, FFG, 
                Habit, FamFFG, FamTolVal, FamHabit) # keep EDAS Master Taxa list names
```

Pull Benthics 

```{r benthics}
benthics <- pool %>% tbl("Edas_Benthic_View") %>%
        as_tibble() %>%
        rename( "StationID" = "STA_ID",
                "BenSampID"  = "WBS_SAMP_ID",
                "RepNum" = "WBS_REP_NUM",
                "FinalID" = "WBMT_FINAL_ID",
                "Individuals" = "WBE_INDIVIDUALS",
                "ID Comments" = "WBE_COMMENT",
                "Entered By" = "WBE_INSERTED_BY", # not in EDAS table but good info
                "Taxonomist" = "TAXONOMIST_NAME",  # not in EDAS table but good info
                "Entered Date" = "WBE_INSERTED_DATE") %>%
        mutate(`Excluded Taxa` = ifelse(WBE_EXCLUDED_TAXA_YN == "Y", -1, 0)) %>%
        dplyr::select(StationID, BenSampID, RepNum, FinalID, Individuals, `Excluded Taxa`, `ID Comments`, Taxonomist, `Entered By`, `Entered Date`)
```

Pull BenSamps 

```{r benSamps}
benSamps <- pool %>% tbl("Edas_Benthic_Sample_View") %>%
        as_tibble() %>%
        # fix names
        rename( "StationID" = "STA_ID",
                "BenSampID"  = "WBS_SAMP_ID",
                "RepNum" = "WBS_REP_NUM",
                "Sample Comments" = "WBS_COMMENT",
                "Entered By" = "WBS_INSERTED_BY", # not in EDAS table but good info
                "Collected By" = "COLLECTOR_NAME",  # not in EDAS table but good info
                "Entered Date" = "WBS_INSERTED_DATE",
                "Gradient" = "WBCM_DESCRIPTION",
                "Taxonomist" = "TAXONOMIST_NAME",  # not in EDAS table but good info
                "Target Count" = "WBS_TARGET_COUNT",
                "Field Team" = "WBS_FIELD_TEAM",
                "Collection Date" = "FDT_DATE_TIME") %>%
        # Add sample season 
        mutate(monthday = as.numeric(paste0(sprintf("%02d",month(`Collection Date`)),
                                            sprintf("%02d",day(`Collection Date`)))),
               Season = case_when(monthday >= 0215 & monthday <= 0615 ~ 'Spring',
                                  monthday >= 0815 & monthday <= 1215 ~ 'Fall',
                                  TRUE ~ as.character("Outside Sample Window"))) %>%
        dplyr::select(StationID, BenSampID, RepNum, `Collection Date`, `Sample Comments`, `Collected By`, `Field Team`, `Entered By`,
                      Taxonomist, `Entered Date`, Gradient, `Target Count`, Season)

```


Pull HabSamps

```{r habSamps}
habSamps <- pool %>% tbl("Edas_Habitat_Sample_View") %>%
        as_tibble() %>%
        rename("StationID" = "STA_ID",
               "HabSampID" = "WHS_SAMP_ID",
               "Entered Date" = "WHS_INSERTED_DATE",
               "Entered By" = "WHS_INSERTED_BY",
               "Field Team" = "WHS_FIELD_TEAM",
               "HabSample Comment" = "WHS_COMMENT",
               "Gradient" = "WSGC_DESCRIPTION",
               "Collection Date" = "FDT_DATE_TIME") %>%
        # Add sample season 
        mutate(monthday = as.numeric(paste0(sprintf("%02d",month(`Collection Date`)),
                                            sprintf("%02d",day(`Collection Date`)))),
               Season = case_when(monthday >= 0215 & monthday <= 0615 ~ 'Spring',
                                  monthday >= 0815 & monthday <= 1215 ~ 'Fall',
                                  TRUE ~ as.character("Outside Sample Window"))) %>%
        dplyr::select(HabSampID, StationID, `Collection Date`, `Entered By`, `Entered Date`, `Field Team`, `HabSample Comment`, Gradient, Season)
      
```

Pull HabValues

```{r}
habValues <- pool %>% tbl("Edas_Habitat_Values_View") %>%
        as_tibble() %>%
        rename("HabSampID" = "WHS_SAMP_ID",
               "HabParameter" = "WHVP_CODE",
               "HabParameterDescription" = "WHVP_DESCRIPTION",
               "HabValue" = "WHV_HAB_VALUE",
               "HabValue Comment" = "WHV_COMMENT") %>%
        dplyr::select(HabSampID, HabParameter, HabParameterDescription, HabValue, `HabValue Comment`)
```

Pull WQM_Stations_View.

```{r}
WQM_Stations_View <- pool %>% tbl("Wqm_Stations_View") %>%  
  as_tibble()
```

Pull WQM_Sta_GIS_View.

```{r}
WQM_Sta_GIS_View <- pool %>% tbl("Wqm_Sta_GIS_View") %>%  
  as_tibble()
```

Pull HabObs

```{r habObs}
habObs <- pool %>% tbl("Edas_Habitat_Observations_View") %>%
        as_tibble() %>%
        rename("HabSampID" = "WHS_SAMP_ID",
               "ObsParameter" = "WOBP_CODE",
               "ObsParameterDescription" = "WOBP_DESCRIPTION",
               "ObsValue" = "WOB_OBS_VALUE") %>%
        dplyr::select(HabSampID, ObsParameter, ObsParameterDescription, ObsValue) 
```




Pull WQM Stations Full From GIS REST service.

```{r wqm_stations_full}
WQM_Station_Full <- suppressWarnings(
      geojson_sf("http://gis.deq.virginia.gov/arcgis/rest/services/staff/DEQInternalDataViewer/MapServer/104/query?&where=OBJECTID%3E0&outFields=*&f=geojson")) %>%
      mutate(WQM_YRS_YEAR = as.Date(as.POSIXct(as.numeric(gsub(" ", ".",gsub("(.{10})", "\\1 ", sprintf("%.0f", as.numeric(WQM_YRS_YEAR))))), origin="1970-01-01")))

WQM_Station_Full <- bind_cols(WQM_Station_Full, st_coordinates(WQM_Station_Full) %>% as.tibble()) %>%
  mutate(Latitude = Y, Longitude = X) %>% dplyr::select(-c(X, Y))

```



Calculate SCI for everyone

```{r sci helpers}
vmast <- masterTaxaGenus %>%
        # get Family level tolerance value, FFG
        rename('GenusTolVal' = 'TolVal',
               'TolVal' = 'FamTolVal',
               'GenusFFG' = 'FFG',
               'FFG' = 'FamFFG',
               'GenusHabit' = 'Habit',
               'Habit' = 'FamHabit') %>%
        mutate(e=ifelse(Order=="Ephemeroptera", 1, 0),
               p=ifelse(Order=="Plecoptera",1,0),
               t=ifelse(Order=="Trichoptera", 1, 0),
               tmin=ifelse((Order=="Trichoptera" & Family != "Hydropsychidae") | 
                             (Order=="Trichoptera" & is.na(Family)) , 1, 0), 
               ept=ifelse(e+p+t>0,1,0), 
               scraper = ifelse(FFG=="Scraper", 1, 0),
               chiro = ifelse(Family=="Chironomidae",1, 0),
               ptmin = ifelse(p + tmin > 0,1,0),
               `clinger-HS` = ifelse(Habit == 'Clinger' & ! Family %in% c("Hydropsychidae","Simuliidae"), 1, 0)) %>%
        # Then put it in long format so it can be merged to and input taxa list
        select(`Final VA Family ID`,TolVal, e,p,t, ept,ptmin, scraper, chiro,`clinger-HS`) %>% 
        distinct(`Final VA Family ID`, .keep_all = T) %>% # drop multiple rows bc working back to family level data from genus
        filter(!is.na(`Final VA Family ID`)) %>%
        pivot_longer(-`Final VA Family ID`, names_to = 'metric', values_to = 'metric_val') %>%
        #  pivot_longer(-`Final VA Family ID`, names_to = 'metric', values_to = 'metric_val') %>%
        filter(!is.na(metric_val))

### SCI calculation

SCI <- function(stationBenthicsDateRange, SCIchoice, benSamps, masterTaxaGenus, vmast){
  edas_options <- select(masterTaxaGenus, Class, Subclass, Order, Suborder, Superfamily, Family, `Final VA Family ID`, FinalID) %>%
    mutate(across(where(is.factor), as.character))
  edas_list <- select(edas_options, `Final VA Family ID`,FinalID)
  # for Excluding taxa, need a list of all Family level designations that may end up as a FinalID
  # these are all unique Family names and the few taxa that are the only 
  GenusNames <- c(unique(edas_options$FinalID)[!is.na(unique(edas_options$FinalID))])
  FamilyNames <- unique(edas_options$Family)[!is.na(unique(edas_options$Family))]
  SuperfamilyNames <- unique(edas_options$Superfamily)[!is.na(unique(edas_options$Superfamily))]
  SuborderNames <- unique(edas_options$Suborder)[!is.na(unique(edas_options$Suborder))]
  OrderNames <- unique(edas_options$Order)[!is.na(unique(edas_options$Order))]
  SubclassNames <- unique(edas_options$Subclass)[!is.na(unique(edas_options$Subclass))]
  ClassNames <- unique(edas_options$Class)[!is.na(unique(edas_options$Class))]
  
  
  EDASrare <- stationBenthicsDateRange %>%
    ########## #filter(str_detect(BenSampID, 'R110') & RepNum == 1) %>% # keep only rarified data and Rep1's
    mutate(Count = Individuals) %>% # Rename to match formatting of functions
    ######`Excluded Taxa` = ifelse(`Excluded Taxa` == T, -1, 0)) %>% 
    select(BenSampID, FinalID, Count, `Excluded Taxa`) %>%
    mutate(GenusTaxaLevel = ifelse(FinalID %in% GenusNames, T, F),
           FamilyTaxaLevel = ifelse(FinalID %in% FamilyNames, T, F),
           SuperfamilyTaxaLevel = ifelse(FinalID %in% SuperfamilyNames, T, F),
           SuborderTaxaLevel = ifelse(FinalID %in% SuborderNames, T, F),
           OrderTaxaLevel = ifelse(FinalID %in% OrderNames, T, F),
           SubclassTaxaLevel = ifelse(FinalID %in% SubclassNames, T, F),
           ClassTaxaLevel = ifelse(FinalID %in% ClassNames, T, F))
  
  # Work FinalID back up to Family Level
  EDASrare2 <- left_join(EDASrare,edas_list, by="FinalID") %>%
    filter(!is.na(`Final VA Family ID`)) %>%
    rename( `Genus Level Excluded Taxa` = `Excluded Taxa`)
  
  # We also need to do a little data manipulation to incorporate biologist exclusion information appropriately.
  exclusionMath  <- EDASrare2 %>%
    mutate(`Family Level Excluded Taxa` = 
             ifelse(`Genus Level Excluded Taxa` == -1, 
                    ifelse(`SuperfamilyTaxaLevel` == TRUE | `SuborderTaxaLevel` == TRUE | `OrderTaxaLevel` == TRUE | 
                             `SubclassTaxaLevel` == TRUE | `ClassTaxaLevel` == TRUE , -1, 0), 0 )) %>%
    # had to get super ifelse nesty here to make this logic work, ugly but works
    group_by(BenSampID, `Final VA Family ID`) %>%
    summarise(`Family Level Count` = sum(Count), 
              #`Genus Level Excluded Taxa` = sum(`Genus Level Excluded Taxa`),
              `Family Level Taxa` = n(),
              `Family Level Excluded Taxa` = sum(`Family Level Excluded Taxa`),
              `Final Family Level Taxa` = `Family Level Taxa` + sum(`Family Level Excluded Taxa`) )
  
  # Join bug traits
  bugTraits <- left_join(exclusionMath,vmast,by=c('Final VA Family ID') )
  
  
  if(SCIchoice == 'VSCI'){SCI <- VSCIcalculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VSCI',
           `SCI Threshold` = 60) %>% 
    rename("SCI Score" ="Fam SCI")}
  if(SCIchoice == 'VCPMI + 63'){SCI <- VCPMI63calculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VCPMI + 63',
           `SCI Threshold` = 42) %>% 
    rename("SCI Score" ="CPMI63+CHOWAN")}
  if(SCIchoice == 'VCPMI - 65'){SCI <- VCPMI65calculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VCPMI - 65',
           `SCI Threshold` = 42) %>% 
    rename("SCI Score" ="CPMI65-CHOWAN")}
  
  SCI <- left_join(SCI, benSamps, by = 'BenSampID')
  
  return(SCI) 
}
```


Calculate VSCI for everyone

```{r vsci}
VSCIresults <- SCI(benthics, "VSCI", benSamps,  masterTaxaGenus, vmast) %>%
        mutate_if(is.numeric, round, digits=2) %>%# rounds all numeric columns to 2 decimal places
        arrange(`Collection Date`)
VSCIresults$Season <-  factor(VSCIresults$Season,levels=c("Spring","Outside Sample Window","Fall"))#,ordered=T)
```



Calculate VCPMI 63 for everyone

```{r vcpmi63}
VCPMI63results <- SCI(benthics, "VCPMI + 63", benSamps,  masterTaxaGenus, vmast) %>%
        mutate_if(is.numeric, round, digits=2) %>%# rounds all numeric columns to 2 decimal places
        arrange(`Collection Date`)
VCPMI63results$Season <-  factor(VCPMI63results$Season,levels=c("Spring","Outside Sample Window","Fall"))#,ordered=T)
```


Calculate VCPMI 65 for everyone

```{r vcpmi65}
VCPMI65results <- SCI(benthics, "VCPMI - 65", benSamps,  masterTaxaGenus, vmast) %>%
        mutate_if(is.numeric, round, digits=2) %>%# rounds all numeric columns to 2 decimal places
        arrange(`Collection Date`)
VCPMI65results$Season <-  factor(VCPMI65results$Season,levels=c("Spring","Outside Sample Window","Fall"))#,ordered=T)
```



Pull Enforcement Data

```{r, eval = FALSE}
# Facilities Data (for joining)
enfFacilities <- pool %>% tbl("Enf_Facilities_View") %>%
  as_tibble()

# get start of year filter information
startOfYear <- as.Date(paste0(year(Sys.Date()), '-01-01'))

# Case data
enfCases <- pool %>% tbl("Enf_enforcement_Cases_View") %>%
  filter(between(Enc_Executed_Date, !! startOfYear, !! Sys.Date()) |
           between(Enc_Nov_Date, !! startOfYear, !! Sys.Date()) |
           between(Enc_Terminated_Date, !! startOfYear, !! Sys.Date()) ) %>%
  as_tibble() %>%
  dplyr::select(Enc_Enf_Facility_Id, 
                `EA Number` = Enc_Ea_Number,
                `Facility Region` = EFC_REGION,
                `Case Manager` = Enf_Case_Mgr,
                `Registration Number` = Enc_Reg_Fac_Num,
                `Program Name` = Program_Name,
                `Permit Number` = Enc_Pmt_Pc_Num,
                `NOV Number` = Enc_Nov_Number,
                `NOV Date` = Enc_Nov_Date,
                `Executed Date` = Enc_Executed_Date,
                `Term Date` = Enc_Terminated_Date, 
                `Cash Civil Charge` = ECC_CASH_CIVIL_CHARGE, 
                `Action Code` = EAC_DESCRIPTION, 
                `Status Code` = Status_Code_Desc, 
                Comments = Enc_Status_Comments) %>%
  left_join(dplyr::select(enfFacilities, Enc_Enf_Facility_Id = Efc_Id, `Facility Name` = Efc_Facility_Name),
            by = "Enc_Enf_Facility_Id") %>%
  dplyr::select(`Facility Name`, everything(), -c(Enc_Enf_Facility_Id)) %>%
  # drop all rows with no data
  drop_na(`Facility Name`, `EA Number`) %>%
  # first fix date fields
  mutate(`NOV Date` = as.POSIXct(`NOV Date`, format = "%m/%d/%Y", tz = 'EST'),
         `Executed Date` = as.POSIXct(`Executed Date`, format = "%m/%d/%Y", tz = 'EST'),
         `Term Date` = as.POSIXct(`Term Date`, format = "%m/%d/%Y", tz = 'EST')) %>%
  # fix factor levels for plotting by program
  mutate(Program = as.factor(case_when(`Program Name` %in% c( "Solid Waste", "Hazardous Waste") ~ "Waste",
                                       `Program Name` %in% c("VWPP") ~ "VWP",
                                       is.na(`Program Name`) | 
                                         `Program Name` %in% c("PReP", "VPA", 
                                                               "Multimedia", "Oil Spill", 
                                                               "Other", "Groundwater","Const. Storm.") ~ "Other", 
                                       `Program Name` %in% c("UST", "AST") ~ "AST/UST", 
                                       TRUE ~ as.character(`Program Name`)))) %>%
  # add date of pull and calculate days Pending
  mutate(pullDate = as.Date(Sys.Date()),
         `Days Pending` = case_when(is.na(`Term Date`) ~ trunc(time_length(interval(`NOV Date`, pullDate), unit = 'day')),
                                    TRUE ~ trunc(time_length(interval(`NOV Date`, `Term Date`), unit = 'day'))))

```




Pin fresh data to board.

```{r pin to rsconnect}
pin(masterTaxaGenus, description = "Edas_Benthic_Master_Taxa_View from ODS", board = "rsconnect")
pin(benthics, description = "Edas_Benthic_View from ODS", board = "rsconnect")
pin(benSamps, description = "Edas_Benthic_Sample_View from ODS", board = "rsconnect")
pin(habSamps, description = "Edas_Habitat_Sample_View from ODS", board = "rsconnect")
pin(habValues, description = "Edas_Habitat_Values_View from ODS", board = "rsconnect")
pin(habObs, description = "Edas_Habitat_Observations_View from ODS", board = "rsconnect")


pin(VSCIresults, description = "VSCI results calculated from ODS_test", board = "rsconnect")
pin(VCPMI63results, description = "VCPMI+63 results calculated from ODS", board = "rsconnect")
pin(VCPMI65results, description = "VCPMI-65 results calculated from ODS", board = "rsconnect")


pin(WQM_Stations_View, description = "Wqm_Stations_View from ODS", board = "rsconnect")
pin(WQM_Sta_GIS_View, description = "Wqm_Sta_GIS_View from ODS", board = "rsconnect")
pin(WQM_Station_Full, description = "first 25k records from WQM Stations (full) from GIS REST service", board = "rsconnect")


#pin(enfFacilities, description = "Enf_Facilities_View from ODS", board = "rsconnect")
#pin(enfCases, description = "Enforcement Cases filtered to start of year", board = "rsconnect")

```






Get data back from rsconnect
```{r reverse}
pin_get("masterTaxaGenus", board = "rsconnect")
```


